% The MIT License (MIT)
%
% Copyright (c) 2020-2021 Yegor Bugayenko
%
% Permission is hereby granted, free of charge, to any person obtaining a copy
% of this software and associated documentation files (the "Software"), to deal
% in the Software without restriction, including without limitation the rights
% to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
% copies of the Software, and to permit persons to whom the Software is
% furnished to do so, subject to the following conditions:
%
% The above copyright notice and this permission notice shall be included
% in all copies or substantial portions of the Software.
%
% THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
% IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
% FITNESS FOR A PARTICULAR PURPOSE AND NON-INFRINGEMENT. IN NO EVENT SHALL THE
% AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
% LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
% OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
% SOFTWARE.

\documentclass[sigplan,nonacm=true]{acmart}
\settopmatter{printfolios=false,printccs=false,printacmref=false}
\usepackage{eo}
\usepackage{iexec}

\acmBooktitle{untitled}
\title{EOLANG and \texorpdfstring{$\varphi$}{phi}-calculus}
\subtitle{\anon{}{%
  git:\iexec{git log -n 1 --pretty='format:\%h'}
  \iexec{git log -n 1 --pretty='format:\%ci'}%
}}
\author{Yegor Bugayenko}
\orcid{0000-0001-6370-0678}
\email{yegor256@gmail.com}
\affiliation{
  \institution{}
  \city{Moscow}
  \country{Russia}
}
\ccsdesc[300]{Software and its engineering~Software notations and tools~Formal language definitions}
\keywords{Object-Oriented Programming, Object Calculus}

\begin{document}

\begin{abstract}
Object-oriented programming (OOP) is one of the most popular
paradigms used for building software systems. However, despite
its industrial and academic popularity, OOP is still missing
a formal apparatus similar to $\lambda$-calculus, which functional
programming is based on. There were a number of attempts to formalize
OOP, but none of them managed to cover all the features available in
modern OO programming languages, such as C++ or Java.
We have made yet another attempt and created \phic{}. We also
created EOLANG (also called \eo{}), an experimental
programming language based on \phic{}.
\end{abstract}

\maketitle

\section{Introduction}
\label{sec:intro}
\input{sections/intro}
\input{sections/overview}

The rest of the paper is dedicated to the discussion of the
syntax of the language we created based on the calculus,
the calculus itself, its semantics, and pragmatics.
In order to make it easier to understand, we start
the discussion with the syntax of the language, while the calculus
is derived from it. Then, we discuss the
key features of \eo{} and the differences between it and other
programming languages. We also discuss how the absence of traditional
OOP features, such as mutability or inheritance, affect the complexity of code.
At the end of the paper we overview the work done by others in the area of
formalization of OOP.

\section{Syntax}
\label{sec:syntax}
\input{sections/syntax}

\section{Calculus}
\label{sec:calculus}
\input{sections/calculus}

\section{Semantics}
\label{sec:semantics}
\input{sections/semantics}

\section{Pragmatics}
\label{sec:pragmatics}
\input{sections/pragmatics}

\section{XMIR}
\label{sec:xmir}
\input{sections/xmir}

\section{Examples}
\label{sec:examples}
\input{sections/examples}

\section{Mappings}
\label{sec:mappings}
\input{sections/mappings}

\section{Key Features}
\label{sec:features}
\input{sections/features}

\section{Four Principles of OOP}
\label{sec:four}
\input{sections/four}

\section{Complexity}
\label{sec:complexity}
\input{sections/complexity}

\section{Related Works}
\label{sec:related}
\input{sections/related}

\section{Acknowledgments}
\input{sections/acks}

\bibliographystyle{ACM-Reference-Format}
\raggedright
\bibliography{main}
\clearpage

\end{document}
